\section{Student Project Proposal}

\subsection{LNU Supervisor}
Oxana Lundström

\subsection{Supervision status according to student}

\begin{itemize}
    \item [x] \textcolor{green}{Currently working with (ongoing project)}
    \item [ ] Agreed, but not started
    \item [ ] Wish to work with (or unclear of status)
    \item [ ] No preference, help me find one
\end{itemize}

\subsection{Cooperative Partners}

I am cooperating with UPTILT, who are currently working on delivering a
program to customers within the service business.
They aim to deliver a product meant to keep track of materials and hourly work
to automate the delivery of invoices and offers to the customer.

I have continuous discussions and meetings with the company.

\subsection{Preliminary Title}

Automated Text Categorization using LLMs for Work Orders

\subsection{Elevator Pitch}

Automatic text categorization is an important step when analyzing textual data,
and the approach of Large Language Models (LLMs) have emerged as an alternative
to the traditional Machine Learning (ML).
ML often requires balanced labeled datasets for effective training
and demands significant manual effort, in contrast to LLMs that
leverages pre-trained contextual knowledge.
I aim to use LLMs to categorize work orders (WOs) titles and descriptions
through a rigorous method,
and evaluate the impact of different prompts and strategies to find
an effective solution to the subject at hand.

\subsection{Steps/Milestones/Actions}

\begin{enumerate}
    \item Find related work.
    \item Define research questions.
    \item Define method.
    \item Conduct experiment.
    \item Elaborate how this method contributes to the field.
    \item Analyze with other studies.
    \item Iterate.
\end{enumerate}

\subsection{Risks}

\begin{itemize}
    \item No previous experience in creating a method, so I will have to rely
          on our supervisor.
    \item Finding relevant previous work related to our subject when it comes
          to using LLMs for text classification.
    \item Limited access certain state-of-the-art models.
\end{itemize}
